---
title: "Sim_ABCRF"
author: "LEJEUNE Thomas"
date: "`r Sys.Date()`"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This R Markdown document is made interactive using Shiny. Unlike the more traditional workflow of creating static reports, you can now create documents that allow your readers to change the assumptions underlying your analysis and see the results immediately. 

To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).

## Inputs and Outputs

You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change.  This demonstrates how a standard R plot can be made interactive by wrapping it in the Shiny `renderPlot` function. The `selectInput` and `sliderInput` functions create the input widgets used to drive the plot.

```{r eruptions, echo=FALSE}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

## Embedded Application

It's also possible to embed an entire Shiny application within an R Markdown document using the `shinyAppDir` function. This example embeds a Shiny application located in another directory:

```{r tabsets, echo=FALSE}
shinyAppDir(
  system.file("examples/06_tabsets", package = "shiny"),
  options = list(
    width = "100%", height = 550
  )
)
```

Note the use of the `height` parameter to determine how much vertical space the embedded application should occupy.

You can also use the `shinyApp` function to define an application inline rather then in an external directory.

In all of R code chunks above the `echo = FALSE` attribute is used. This is to prevent the R code within the chunk from rendering in the document alongside the Shiny components.





```{r = Packages generation} 

library(tidyverse)
library(abcrf)
library(ggplot2)

```

```{r = Get the data}

# Load the csv file from the "simulations" directory
data <- read_csv("Bureau/simulations/Ref_table/summary_table.csv")

```

```{r = Hard cleaning (Column removal - Automatic version)}

# Function that will search every column and count the "NA", then remove the columns having too much missing values according to the threshold
remove_high_na_cols <- function(data, threshold = 0.2) {
  na_prop <- colSums(is.na(data)) / nrow(data)
  cols_to_remove <- names(na_prop[na_prop > threshold])
  data_clean <- data %>% select(-any_of(cols_to_remove))
  list(data = data_clean, removed = cols_to_remove)
}
result <- remove_high_na_cols(data, threshold = 0.2)
data <- result$data 
result$removed  # List of removed col

```

```{r = Soft cleaning (Median)}

# Search every column for NA then calculate the median of the existing values and add it to the missing values
data_imputed <- data %>% mutate(across(everything(), ~ ifelse(is.na(.), median(., na.rm = TRUE), .)))

```

```{r = Hard cleaning (column removal - Classic version)}

# Hard version of the last method, choose the columns to delete to avoid suppressing too many lines while removing the NAs
to_remove <- c(
  "HE_Neb_mean_0.050_Pop2", "HE_Neb_mean_0.020_Pop2", "HE_Neb_mean_0.010_Pop2", "HE_Neb_mean_0.000_Pop2",
  "HE_Neb_mean_0.050_Pop1", "HE_Neb_mean_0.020_Pop1", "HE_Neb_mean_0.010_Pop1", "HE_Neb_mean_0.000_Pop1",
  "Coan_Neb_n_Pop1", "Coan_Neb_n_Pop2",
  "LD_Ne_0.050_Pop1", "LD_Ne_0.020_Pop1", "LD_Ne_0.010_Pop1", "LD_Ne_0.000_Pop1",
  "LD_Ne_0.050_Pop2", "LD_Ne_0.020_Pop2", "LD_Ne_0.010_Pop2", "LD_Ne_0.000_Pop2"
)

# Remove the lines that contain at least one "NA"
data <- data %>% select(-any_of(to_remove)) %>% na.omit()

```

```{r = Set the parameters and resuming statistics tables}

# List of parameters
param_cols <- c("simulation_id", "pop_size", "num_loci", "sample1_size_Ne", "sample2_size_Ne",
                "sample1_size_CMR", "sample2_size_CMR", "mutation_rate", "recap_Ne")
params <- data %>% select(all_of(param_cols))

# List of resuming statistics
stat_keywords <- c("id", "LD", "HE", "Coan", "het", "alleles", "P_", "N_", "J_")
stat_cols <- names(data)[sapply(names(data), function(col) any(str_detect(col, stat_keywords)))]
stats_table <- data %>% select(all_of(stat_cols))

```

```{r = ABCRF model}

# Choose the parameter that will be predicted and create a new dataframe with it in the first column and the resuming stats
target_param <- "pop_size"
learning_data <- bind_cols(y = params[[target_param]], data_imputed) %>% rename(!!target_param := y)

# Regression ABCRF that will predict the values of a parameters from the data base formed by the simulations
model_rf <- regAbcrf(as.formula(paste(target_param, "~ .")), data = learning_data, ntree = 500)
summary(model_rf)

```

```{r = Stat importance for ABCRF model}

# Determines the order of importance of every resuming stat in the dataframe concerning the prediction of the parameter
importance_df <- model_rf$model.rf$variable.importance %>%
  sort(decreasing = TRUE) %>%
  enframe(name = "Statistic", value = "Importance")
top_n <- 20
importance_top <- importance_df %>%
  slice_head(n = top_n) %>%
  mutate(Statistic = fct_reorder(Statistic, Importance))

# Plotting the barplot showing the order of importance
ggplot(importance_top, aes(x = Statistic, y = Importance)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(
    title = paste("Top", top_n, "variables importantes pour l'estimation de", target_param),
    x = "Statistique",
    y = "Importance (abcrf)"
  ) +
  theme_minimal(base_size = 14)

```

```{r = Prediction values of the model}

# Use the model to make the predictions according to the "observed data" from the Out-Of-Bag "OOB"
predictions <- predict(model_rf, training = learning_data, obs = learning_data)$expectation

# Plotting the curve demonstrating the differences of the predictions of the dataframe studied with the target data from the Out-Of-Bag
plot(learning_data[[target_param]], predictions,
     xlab = "Vraie pop_size", ylab = "Prédiction pop_size (OOB)",
     main = "Comparaison vraie vs prédite",
     pch = 20, col = "blue")
abline(0, 1, col = "red")

```

